ICM testing

### 2023-10-03 Scripts/Paper_1.Rmd
Forward and backward tuning with fecundity working

```{r, load-stuff,echo=F, message=F, warning=F, fig.width=6, fig.height=4}
# OK, so using the ICES assessments here's what we get for North Sea cod.
library(readxl)
library(tidyverse)
library(rio)
library(ggthemes)
library(cowplot)
loc <- 'C:/Users/Owner/Documents/GitHub/ICM'

@@

back.res <- NULL
back.ts <- NULL
back.ts.quants <- NULL
for.res <- NULL
for.ts <- NULL
for.ts.quants <- NULL
tuned.res <- NULL
for(i in Stocks)
{
  years <- years.tmp[[i]]
  prop.nat.mort <- pnm.tmp[[i]] 
  weight.age <- waa.tmp[[i]] 
  ages <- ages.tmp[[i]] 
  rem <- rem.tmp[[i]] 
  mx <- mx.tmp[[i]] 
  N.end <- NE.tmp[[i]] 
  vpa.abund <- vpa.tmp[[i]] 
  age.mat <- am.tmp[[i]]

  tst <- back.sim(years,
                 mat.age = age.mat,
                 nm = -(log(1-prop.nat.mort)),
                 w.age = weight.age,
                 ages = ages,
                 rems = rem$rem,
                 fecund = mx,
                 N.end = N.end,
                 pop.model = 'exponential',
                 n.sims = 50,
                 sd.mat = 0.5,
                 sd.nm = 0.5,
                 sd.wt = 0.5,
                 sd.fecund = 0.5)

@@

# ################# Now run as a forward simulation
#   N.start <- vpa.abund[1] 
# 
#   tst <- for.sim(years,
#                  mat.age = age.mat,
#                  nm = -(log(1-prop.nat.mort)),
#                  w.age = weight.age,
#                  ages = ages,
#                  rems = rem$rem, # rep(0,nrow(rem)),
#                  fecund = mx,
#                  N.start = N.start,
#                  pop.model = 'exponential', K = 2e6,
#                  n.sims = 2,
#                  sd.mat = 0.5,
#                  sd.nm = 0.5,
#                  sd.wt = 0.5,
#                  sd.fecund = 0.5)
# 
#   #Combine the data
#   did.it.work <- data.frame(abund = c(vpa.abund,tst$Pop$abund),years = c(years,tst$Pop$years),sim = c(rep('VPA',length(years)),tst$Pop$sim))
#   # Here get the Upper and lower 50% quantiles to make a functional boxplot
#   quants <- did.it.work %>% dplyr::filter(sim != "VPA") %>% dplyr::group_by(years) %>% dplyr::summarise(L.50 = quantile(abund,probs=c(0.25)),
#                                                                         med = median(abund),
#                                                                         U.50 = quantile(abund,probs=c(0.75)))
# 
#   for.res[[i]] <- tst
#   for.ts[[i]] <- did.it.work
#   for.ts.quants[[i]] <- quants

# ########################## Now run the tuning sims
# #
#    N.end <- vpa.abund[length(vpa.abund)]
#    tunes <- tune.sim(years,
#                      n.steps = 100,
#                      abund.ts = vpa.abund,
#                      mat.age = age.mat,
#                      nm = -(log(1-prop.nat.mort)),
#                      w.age = weight.age,
#                      ages = ages,
#                      rems = rem$rem,
#                      fecund = mx,
#                      N.end = N.end,
#                      pop.model = 'exponential')
# 
#    tuned.res[[i]] <- tunes
} # end i in Stocks look

# Save all these
# saveRDS(object = tuned.res,file = "D:/Github/ICM/Results/tuned_fits.Rds")
 saveRDS(object = back.res,file = paste0(loc,"/Results/backwards_sim_results.Rds"))
 saveRDS(object = back.ts,file = paste0(loc,"/Results/backwards_sim_ts_summaries.Rds"))
 saveRDS(object = back.ts.quants,file = paste0(loc,"/Results/backwards_sim_ts_quantiles_summaries.Rds"))
```

```{r, make-figs,echo=F, message=F, warning=F, fig.width=6, fig.height=4}

# Now we can run the figures separately.
tuned.res <- readRDS(file = "D:/Github/ICM/Results/tuned_fits.Rds")
back.ts <- readRDS(file = "D:/Github/ICM/Results/backwards_sim_ts_summaries.Rds")
back.ts.quants <- readRDS(file = "D:/Github/ICM/Results/backwards_sim_ts_quantiles_summaries.Rds")
back.res <- readRDS(file = "D:/Github/ICM/Results/backwards_sim_results.Rds")
load(file = "D:/Github/ICM/Results/model_inputs.Rdata")

for(i in Stocks)
{
  # Unpack the lists
  tunes <- tuned.res[[i]] 
  did.it.work <- back.ts[[i]] 
  quants <- back.ts.quants[[i]] 
  missing_rem <- mr.tmp[[i]] 
  tst <- back.res[[i]]

  did.it.work$missing[did.it.work$years %in% missing_rem] <- "missing"

    # All the results in a plot
  p2 <- ggplot() +
          geom_line(data=did.it.work, aes(x=years,y=abund,color=sim)) +
          xlab("") +
          ylab("Abundance (1000s)") +
          geom_line(data=did.it.work %>% dplyr::filter(sim == "VPA"),aes(x=years,y=abund),color='black',size=2) +
          #scale_y_continuous(breaks = seq(0,3e6,by=5e5)) + scale_x_continuous(breaks = seq(1960,2025,by=5)) +
          theme_few() + theme(legend.position = 'none') + scale_color_viridis_d(end = 0.75) +
          ggtitle(i) +
          geom_text(data=did.it.work[did.it.work$missing=="missing",], aes(x=years, y=abund, label="?"))
  save_plot(plot = p2, filename = paste0("D:/Github/ICM/Figures/raw/", i, "/runs_vs_vpa_", i, ".png"), base_width = 6,base_height = 4)
  # Same thing but functional boxplots
  p3 <- ggplot() +
          geom_line(data=quants, aes(x=years,y= med)) +
          geom_ribbon(data=quants, aes(x=years,ymax=U.50,ymin = L.50),alpha=0.5,fill='blue',color='blue') +
          geom_line(data=did.it.work %>% dplyr::filter(sim == "VPA"),aes(x=years,y=abund),color='black',size=2) +
          xlab("") + ylab("Abundance (1000s)") +
          #scale_y_continuous(breaks = seq(0,3e6,by=5e5)) + scale_x_continuous(breaks = seq(1960,2025,by=5)) +
          theme_few() + theme(legend.position = 'none') + #geom_vline()
          ggtitle(i)+
          geom_text(data=did.it.work[did.it.work$missing=="missing",], aes(x=years, y=abund, label="?"))
  save_plot(plot = p3, filename = paste0("D:/Github/ICM/Figures/raw/", i, "/runs_vs_vpa_ribbon_", i, ".png"), base_width = 6,base_height = 4)

  # Plotting how the estimate of r changes over time (which you can do when we have time varying inputs (e.g. Wgt, M, or maturity at age)
  p4 <- ggplot(tst$r) +
          geom_line(aes(x=years,y=r,group=n.sims,color=n.sims)) +
          xlab("") +
          ylab("Lotka r") +
          #scale_y_continuous(breaks = seq(0,1.5,by=0.1)) +
          #scale_x_continuous(breaks = seq(1960,2025,by=5)) +
          theme_few() + theme(legend.position = 'none') +
          scale_color_viridis_c(end = 0.75) +
          ggtitle(i)
  save_plot(plot = p4, filename = paste0("D:/Github/ICM/Figures/raw/", i, "/lokta_r_", i, ".png"), base_width = 6,base_height = 4)

  # Comparing our new time series with the old
  p2 <- ggplot() +
          geom_line(data=tunes$res, aes(x=year,y=lotka.n),color='blue') +
          geom_line(data=tunes$res, aes(x=year,y=vpa.n),color='black') +
          xlab("") +
          ylab("Abundance (1000s)") +
          #geom_line(data=did.it.work %>% dplyr::filter(sim == "VPA"),aes(x=years,y=abund),color='black',size=2) +
          #scale_y_continuous(breaks = seq(0,3e6,by=5e5)) + scale_x_continuous(breaks = seq(1960,2025,by=5)) +
          theme_few() + theme(legend.position = 'none') +
          ggtitle(i) #+
          #geom_text(data=did.it.work[did.it.work$missing=="missing",], aes(x=years, y=abund, label="?"))
  save_plot(plot = p2, filename = paste0("D:/Github/ICM/Figures/tuned/", i, "/tunded_vs_vpa_", i, ".png"), base_width = 6,base_height = 4)

  # Comparing the absolute miss from the tuned time series
   p3 <- ggplot() +
          #geom_line(data=tunes$res, aes(x=year,y=lotka.n),color='blue') +
          geom_line(data=tunes$res[-1,], aes(x=year,y=diff.n),color='red') +
          xlab("") +
          ylab("Difference in abundance (1000s)") +
          #geom_line(data=did.it.work %>% dplyr::filter(sim == "VPA"),aes(x=years,y=abund),color='black',size=2) +
          #scale_y_continuous(breaks = seq(0,3e6,by=5e5)) + scale_x_continuous(breaks = seq(1960,2025,by=5)) +
          theme_few() + theme(legend.position = 'none') + 
          ggtitle(i) 
          #geom_text(data=did.it.work[did.it.work$missing=="missing",], aes(x=years, y=abund, label="?"))
  save_plot(plot = p3, filename = paste0("D:/Github/ICM/Figures/tuned/", i, "/diff_N_", i, ".png"), base_width = 6,base_height = 4)

  # What percentage (mean) difference
   p4 <- ggplot() +
          #geom_line(data=tunes$res, aes(x=year,y=lotka.n),color='blue') +
          geom_line(data=tunes$res[-1,], aes(x=year,y=per.diff.n),color='red') +
          xlab("") +
          ylab("Difference in abundance (%)") +
          #geom_line(data=did.it.work %>% dplyr::filter(sim == "VPA"),aes(x=years,y=abund),color='black',size=2) +
          #scale_y_continuous(breaks = seq(0,3e6,by=5e5)) + scale_x_continuous(breaks = seq(1960,2025,by=5)) +
          theme_few() + theme(legend.position = 'none') + 
          ggtitle(i) #+ ylim(c(-100,100))
          #geom_text(data=did.it.work[did.it.work$missing=="missing",], aes(x=years, y=abund, label="?"))
  save_plot(plot = p4, filename = paste0("D:/Github/ICM/Figures/tuned/", i, "/per_diff_N_", i, ".png"), base_width = 6,base_height = 4)

     # Difference in m's
   p5 <- ggplot() +
          geom_line(data=tunes$res[-1,], aes(x=year,y=diff.m),color='red') +
          xlab("") +
          ylab("Difference in natural mortality (Instantaneous)") +
          #geom_line(data=did.it.work %>% dplyr::filter(sim == "VPA"),aes(x=years,y=abund),color='black',size=2) +
          #scale_y_continuous(breaks = seq(0,3e6,by=5e5)) + scale_x_continuous(breaks = seq(1960,2025,by=5)) +
          theme_few() + theme(legend.position = 'none') + 
          ggtitle(i) #+ ylim(c(-100,100))
          #geom_text(data=did.it.work[did.it.work$missing=="missing",], aes(x=years, y=abund, label="?"))
  save_plot(plot = p5, filename = paste0("D:/Github/ICM/Figures/tuned/", i, "/diff_m_", i, ".png"), base_width = 6,base_height = 4)
     # % Difference in m's
   p6 <- ggplot() +
          geom_line(data=tunes$res[-1,], aes(x=year,y=per.diff.m),color='red') +
          xlab("") +
          ylab("Difference in natural mortality (%)") +
          #geom_line(data=did.it.work %>% dplyr::filter(sim == "VPA"),aes(x=years,y=abund),color='black',size=2) +
          #scale_y_continuous(breaks = seq(0,3e6,by=5e5)) + scale_x_continuous(breaks = seq(1960,2025,by=5)) +
          theme_few() + theme(legend.position = 'none') + 
          ggtitle(i) #+ ylim(c(-100,100))
          #geom_text(data=did.it.work[did.it.work$missing=="missing",], aes(x=years, y=abund, label="?"))
  save_plot(plot = p6, filename = paste0("D:/Github/ICM/Figures/tuned/", i, "/per_diff_m_", i, ".png"), base_width = 6,base_height = 4)

   p7 <- ggplot() +
          geom_line(data=tunes$res[-1,], aes(x=year,y=mn.m),color='blue') +
          geom_line(data=tunes$res[-1,], aes(x=year,y=vpa.m),color='black') +
          xlab("") +
          ylab("Natural mortality (Instantaneous)") +
          #geom_line(data=did.it.work %>% dplyr::filter(sim == "VPA"),aes(x=years,y=abund),color='black',size=2) +
          #scale_y_continuous(breaks = seq(0,3e6,by=5e5)) + scale_x_continuous(breaks = seq(1960,2025,by=5)) +
          theme_few() + theme(legend.position = 'none') +
          ggtitle(i) #+
          #geom_text(data=did.it.work[did.it.work$missing=="missing",], aes(x=years, y=abund, label="?"))
  save_plot(plot = p7, filename = paste0("D:/Github/ICM/Figures/tuned/", i, "/Nat_M_", i, ".png"), base_width = 6,base_height = 4)

     p8 <- ggplot() +
          geom_line(data=tunes$res[-1,], aes(x=year,y=r),color='blue') +
          xlab("") +
          ylab("Lotka r") +
          #geom_line(data=did.it.work %>% dplyr::filter(sim == "VPA"),aes(x=years,y=abund),color='black',size=2) +
          #scale_y_continuous(breaks = seq(0,3e6,by=5e5)) + scale_x_continuous(breaks = seq(1960,2025,by=5)) +
          theme_few() + theme(legend.position = 'none') +
          ggtitle(i) #+
          #geom_text(data=did.it.work[did.it.work$missing=="missing",], aes(x=years, y=abund, label="?"))
  save_plot(plot = p8, filename = paste0("D:/Github/ICM/Figures/tuned/", i, "/Lotka_r_", i, ".png"), base_width = 6,base_height = 4)

@@

# summary plots for ALL stocks, this is already run, so lets not keep running it
# for (i in unique(ASR_long$Stock)) {
#   ASR_sub <- ASR_long %>%
#     dplyr::select(Year, Stock, type, age, value) %>%
#     dplyr::filter(Stock==i)
#   maxage <- max(as.numeric(ASR_sub$age[!is.na(ASR_sub$value)]),na.rm=T)
#   maxYear <- max(ASR_sub$Year[!is.na(ASR_sub$value)],na.rm=T)
#   minYear <- min(ASR_sub$Year[!is.na(ASR_sub$value)],na.rm=T)
#   forplot <- ASR_sub[as.numeric(ASR_sub$age) < (maxage+1) & ASR_sub$Year %in% minYear:maxYear,]
#   forplot <- forplot[!is.na(forplot$value),] # remove the NAs, hopefully this works for all the stocks...
#   p1 <- ggplot() + geom_line(data=forplot, aes(Year, value)) + facet_grid(type~as.numeric(age), scales="free_y") +
#         geom_vline(data=forplot[is.na(forplot$value),], aes(xintercept=Year), colour="red") + facet_grid(type~as.numeric(age), scales="free_y") +
#         theme_bw() + 
#         ggtitle(i) +
#     guides(x =  guide_axis(angle = 90)) 
#   print(p1)
#   
#   if(!dir.exists(paste0("D:/Github/ICM/Figures/", i))) dir.create(paste0("D:/Github/ICM/Figures/", i))
#   
#   png(filename=paste0("D:/Github/ICM/Figures/", i, "/data_summary_", i, ".png"), height=4, width=12, res=400, units="in")
#   print(p1)
#   dev.off()
# }

@@

```{r, forward-sims-n-figs,echo=F, message=F, warning=F, fig.width=6, fig.height=4}
# Now we have to forward project the stocks, I think it makes the most sense to do this using the tuned data at the moment
# So we assume the fecundity data and the tuned natural mortality data for the stocks is a reasonable representation of the future
load(file = "D:/Github/ICM/Results/model_inputs.Rdata")

back.res <- NULL
back.ts <- NULL
back.ts.quants <- NULL
for.res <- NULL
for.ts <- NULL
for.ts.quants <- NULL
tuned.res <- NULL
for(i in Stocks)
{
  years <- years.tmp[[i]]
  prop.nat.mort <- pnm.tmp[[i]] 
  weight.age <- waa.tmp[[i]] 
  ages <- ages.tmp[[i]] 
  rem <- rem.tmp[[i]] 
  mx <- mx.tmp[[i]] 
  N.end <- NE.tmp[[i]] 
  vpa.abund <- vpa.tmp[[i]] 
  age.mat <- am.tmp[[i]]

################# Now run as a forward simulation
  N.start <- vpa.abund[length(vpa.abund)] 
  years <- 2025:2075
  # For the forward simulations I'm setting it up so we feed in the mean vector or a character, we'll then create a vector of mx's with variablitity
  st.time <- Sys.time()
  tst <- for.sim(years,
                 mat.age = age.mat,
                 nm = -(log(1-prop.nat.mort)),
                 w.age = weight.age,
                 ages = ages,
                 rems = c(0.1,0.2), # This is set up as a fishing mortality that averages 0.1 with an sd of 0.2)
                 fecund = mx,
                 N.start = N.start,
                 pop.model = 'exponential', sim = 'project', proj.sim = 'sample',
                 n.sims = 30,
                 sd.mat = 0.1,
                 sd.nm = 0.1,
                 sd.wt = 0.1,
                 sd.fecund = 0.1)
  Sys.time() - st.time
  #Combine the data
  did.it.work <- data.frame(abund = tst$Pop$abund,years = c(tst$Pop$years),sim = tst$Pop$sim)
  # Here get the Upper and lower 50% quantiles to make a functional boxplot
  quants <- tst$Pop %>% dplyr::group_by(years) %>% dplyr::summarise(L.50 = quantile(abund,probs=c(0.25)),
                                                                   med = median(abund),
                                                                   U.50 = quantile(abund,probs=c(0.75)))

  for.res[[i]] <- tst
  for.ts[[i]] <- tst$Pop
  for.ts.quants[[i]] <- quants  
}  
ggplot(quants) + geom_line(aes(x=years,y=med))  + geom_ribbon(aes(x=years,ymax=U.50,ymin = L.50),alpha=0.5,fill='blue',color='blue')

>>>>

```{r, load-stuff,echo=F, message=F, warning=F, fig.width=6, fig.height=4}
# OK, so using the ICES assessments here's what we get for North Sea cod.
library(readxl)
library(tidyverse)
library(rio)
library(ggthemes)
library(cowplot)
loc <- 'D:/GitHub/ICM'

@@

back.res <- NULL
back.ts <- NULL
back.ts.quants <- NULL
for.res <- NULL
for.ts <- NULL
for.ts.quants <- NULL
tuned.res <- NULL
for(i in Stocks)
{
  years <- years.tmp[[i]]
  prop.nat.mort <- pnm.tmp[[i]] 
  weight.age <- waa.tmp[[i]] 
  ages <- ages.tmp[[i]] 
  rem <- rem.tmp[[i]] 
  mx <- mx.tmp[[i]] 
  N.end <- NE.tmp[[i]] 
  vpa.abund <- vpa.tmp[[i]] 
  age.mat <- am.tmp[[i]]
  # 
  # # The first run gives us the mean estimate, I think that's all we need for our first paper
  tst <- back.sim(years,
                 mat.age = age.mat,
                 nm = -(log(1-prop.nat.mort)),
                 w.age = weight.age,
                 ages = ages,
                 rems = rem$rem,
                 fecund = mx,
                 N.end = N.end,
                 pop.model = 'exponential',
                 n.sims = 1)

@@

# ################# Now run as a forward simulation
N.start <- vpa.abund[1]

tst <- for.sim(years,
               mat.age = age.mat,
               nm = -(log(1-prop.nat.mort)),
               w.age = weight.age,
               ages = ages,
               rems = rem$rem, # rep(0,nrow(rem)),
               fecund = mx,
               N.start = N.start,
               pop.model = 'exponential',
               n.sims = 1,
               sd.mat = 0,
               sd.nm = 0,
               sd.wt = 0,
               sd.fecund = 0)

  for.res[[i]] <- tst
  for.ts[[i]] <- did.it.work
  for.ts.quants[[i]] <- quants

} # end i in Stocks look

# Save all these
saveRDS(object = back.res,file = paste0(loc,"/Results/backwards_sim_results_no_error.Rds"))
saveRDS(object = back.ts,file = paste0(loc,"/Results/backwards_sim_ts_summaries_no_error.Rds"))
saveRDS(object = back.ts.quants,file = paste0(loc,"/Results/backwards_sim_ts_quantiles_summaries_no_error.Rds"))
# Forward simulations, just using the 'realization'
saveRDS(object = for.res,file = paste0(loc,"/Results/forwards_sim_results_no_error.Rds"))
saveRDS(object = for.ts,file = paste0(loc,"/Results/forwards_sim_ts_summaries_no_error.Rds"))
saveRDS(object = for.ts.quants,file = paste0(loc,"/Results/forwards_sim_ts_quantiles_summaries_no_error.Rds"))

```{r}

# Upack the back.res object...

tst.for <- NULL
tst.back <- NULL
count = 0
for(i in Stocks)
{
  count=count+1
  tmp.for <- for.res[[i]]$r %>% dplyr::filter(n.sims==1)
  tmp.for <- data.frame(tmp.for,stock = i,direction = 'forward simulations',id = count)
  tmp.pop.for <- for.res[[i]]$Pop %>% dplyr::filter(sim==1)
  tst.for[[i]] <- left_join(tmp.pop.for,tmp.for,by=c('years'))
  # backwards
  tmp.back <- back.res[[i]]$r %>% dplyr::filter(n.sims==1)
  tmp.back <- data.frame(tmp.back,stock = i,direction = 'backwards simulations',id = count)
  tmp.pop.back <- back.res[[i]]$Pop %>% dplyr::filter(sim==1)
  tst.back[[i]] <- left_join(tmp.pop.back,tmp.back,by=c('years'))
}

fors <- do.call('rbind',tst.for)
fors <- fors[,-4]
backs <- do.call('rbind',tst.back)

exploring <- rbind(fors,backs)
mn.r <- exploring %>% dplyr::group_by(stock,direction,id) %>% dplyr::summarise(r = median(r,na.rm=T),
                                                                  pop.ratio = dplyr::last(abund)/dplyr::first(abund))
windows(11,11)
ggplot(mn.r) + geom_text(aes(r,pop.ratio,label=id)) + scale_y_log10() + facet_wrap(~direction) + xlab("mean instrinc rate of growth (r)") + ylab("N (Final) / N (Initial)")

quantile(mn.r$r,probs = seq(0.01,1,by=0.01),na.rm=T)

So now I can try to do some tuning, both forwards and backwards, lets see what we get from that...

```{r tunes}

########################## Now run the tuning sims
#

back.res <- NULL
back.ts <- NULL
back.ts.quants <- NULL
for.res <- NULL
for.ts <- NULL
for.ts.quants <- NULL
tuned.res <- NULL
for(i in Stocks)
{
  years <- years.tmp[[i]]
  prop.nat.mort <- pnm.tmp[[i]] 
  weight.age <- waa.tmp[[i]] 
  ages <- ages.tmp[[i]] 
  rem <- rem.tmp[[i]] 
  mx <- mx.tmp[[i]] 
  N.end <- NE.tmp[[i]] 
  vpa.abund <- vpa.tmp[[i]] 
  age.mat <- am.tmp[[i]]
  N.end <- vpa.abund[length(vpa.abund)]

   back.tunes <- tune.sim(years,
                          tuner = 'f',
                           n.steps = 100, step.size = 0.2,
                           abund.ts = vpa.abund,
                           mat.age = age.mat,
                           nm = -(log(1-prop.nat.mort)),
                           w.age = weight.age,
                           ages = ages,
                           rems = rem$rem,
                           fecund = mx,
                           N.init = N.end,
                           pop.model = 'exponential',
                           direction='backwards')

   back.tuned.res[[i]] <- back.tunes

   N.start <- vpa.abund[1]

   for.tunes   <- tune.sim(years,
                           tuner='f',
                           n.steps =100, step.size =0.2,
                           abund.ts = vpa.abund,
                           mat.age = age.mat,
                           nm = -(log(1-prop.nat.mort)),
                           w.age = weight.age,
                           ages = ages,
                           rems = rem$rem,
                           fecund = mx,
                           N.init = N.start,
                           pop.model = 'exponential',
                           direction= 'forwards')

   for.tuned.res[[i]] <- for.tunes

@@

@@
